---
title: "Dataset example - function for wide-to-long format variables"
author: "Miruna Barbu"
date: "28/03/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Read in relevant R packages
library(tidyverse)

```

The accompanying .r script ("wide_to_long_format.R") uses a function created by myself to turn multiple variables from wide to long format. This is useful in analyses that use repeated measures, such as linear mixed-effects models.

The script:

1. Loads in a dummy dataset. In this case, I created a dataset containing measures for three white matter tracts (anterior, posterior, and superior thalamic radiations), for the left and right hemisphere. This is a measure I regularly work with in my research, although the script can be applied to any variable needing to be transformed from wide to long format.

*Although the dummy dataset contains 10 variables only, the script works with large-scale datasets (I have used this with >30,000 participants and up to 64 brain regions).

2. Loads in a covariate file (this can be excluded from the function by commenting out lines ...). This saves time from including covariates at later stages of the analysis process.

3. Carries out a number of steps to prepare the data for input into the function. The function takes 6 arguments (or 5 in case you want to exclude covariates).

The script for creating dummy data is below:

```{r Create dummy data, echo=FALSE}
# Create dummy dataset file

dummy_whitematter <- data.frame(ID = seq(1:10),
                  lh.antthalrad = sample(seq(0.2, 0.5, by= 0.001),10),
                  rh.antthalrad = sample(seq(0.2, 0.5, by= 0.001),10),
                  lh.postthalrad = sample(seq(0.2, 0.5, by= 0.001),10),
                  rh.postthalrad = sample(seq(0.2, 0.5, by= 0.001),10),
                  lh.supthalrad = sample(seq(0.2, 0.5, by= 0.001),10),
                  rh.supthalrad = sample(seq(0.2, 0.5, by= 0.001),10))

# Create dummy covariate file
dummy_cov <- data.frame(ID = seq(1:10),
                        sex=as.factor(sample(0:1, 10, replace = TRUE)),
                        age=sample(seq(25, 40),10))

# Save both
# saveRDS(dummy_whitematter, "./wide_to_long_dummy_dataset.rds")

# saveRDS(dummy_cov, "./wide_to_long_dummy_covfile.rds")



cat("Print summaries of both datasets (for visualisation purposes)")

cat("Dummy dataset summary:")

print(summary(dummy_whitematter))

cat("Dummy covariate summary:")

print(summary(dummy_cov))


```




